---
title: "Dye Sankey Diagram"
output: html_notebook
---

This is an R notebook dedicated to figuring out the best way to make a Sankey diagram for the dye paper. 

Will explore at least: 
- networkD3 (typical sankey package)
- ggplot with geom_smooth and scale_color_viridis_d 

```{r}
library(networkD3)
library(tidyverse)
```
Read in the relevant data
```{r}
# hierarchically clustered dye library, with an order of where the dyes fall in this arrangement
# classification at the following steps:
# 1: initial index group (A, C, MWA, L, etc.)
# 2: +/- detection of protein unfolding
# 3: final library, temp sensitive, reactive, etc. 

```

```{r}
#### this is the sankeyNetwork src from https://github.com/christophergandrud/networkD3/blob/master/R/sankeyNetwork.R
#' # Plot
#' sankeyNetwork(Links = energy$links, Nodes = energy$nodes, Source = 'source',
#'              Target = 'target', Value = 'value', NodeID = 'name',
#'              units = 'TWh', fontSize = 12, nodeWidth = 30)
#'
#' # Colour links
#' energy$links$energy_type <- sub(' .*', '',
#'                                energy$nodes[energy$links$source + 1, 'name'])
#'
#' sankeyNetwork(Links = energy$links, Nodes = energy$nodes, Source = 'source',
#'              Target = 'target', Value = 'value', NodeID = 'name',
#'              LinkGroup = 'energy_type', NodeGroup = NULL)
#'
#' }
#' @source
#' D3.js was created by Michael Bostock. See \url{http://d3js.org/} and, more
#' specifically for Sankey diagrams \url{http://bost.ocks.org/mike/sankey/}.
#'
#' @seealso \code{\link{JS}}
#'
#' @export

sankeyNetwork <- function(Links, Nodes, Source, Target, Value,
    NodeID, NodeGroup = NodeID, LinkGroup = NULL, units = "",
    colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"), fontSize = 7,
    fontFamily = NULL, nodeWidth = 15, nodePadding = 10, margin = NULL,
    height = NULL, width = NULL, iterations = 32, sinksRight = TRUE)
{
    # Check if data is zero indexed
    check_zero(Links[, Source], Links[, Target])

    # Hack for UI consistency. Think of improving.
    colourScale <- as.character(colourScale)

    # If tbl_df convert to plain data.frame
    Links <- tbl_df_strip(Links)
    Nodes <- tbl_df_strip(Nodes)

    # Subset data frames for network graph
    if (!is.data.frame(Links)) {
        stop("Links must be a data frame class object.")
    }
    if (!is.data.frame(Nodes)) {
        stop("Nodes must be a data frame class object.")
    }
    # if Source or Target are missing assume Source is the first
    # column Target is the second column
    if (missing(Source))
        Source = 1
    if (missing(Target))
        Target = 2

    if (missing(Value)) {
        LinksDF <- data.frame(Links[, Source], Links[, Target])
        names(LinksDF) <- c("source", "target")
    } else if (!missing(Value)) {
        LinksDF <- data.frame(Links[, Source], Links[, Target],
            Links[, Value])
        names(LinksDF) <- c("source", "target", "value")
    }

    # if NodeID is missing assume NodeID is the first column
    if (missing(NodeID))
        NodeID = 1
    NodesDF <- data.frame(Nodes[, NodeID])
    names(NodesDF) <- c("name")

    # add node group if specified
    if (is.character(NodeGroup)) {
        NodesDF$group <- Nodes[, NodeGroup]
    }

    if (is.character(LinkGroup)) {
        LinksDF$group <- Links[, LinkGroup]
    }

    margin <- margin_handler(margin)

    # create options
    options = list(NodeID = NodeID, NodeGroup = NodeGroup, LinkGroup = LinkGroup,
        colourScale = colourScale, fontSize = fontSize, fontFamily = fontFamily,
        nodeWidth = nodeWidth, nodePadding = nodePadding, units = units,
        margin = margin, iterations = iterations, sinksRight = sinksRight)

    # create widget
    htmlwidgets::createWidget(name = "sankeyNetwork", x = list(links = LinksDF,
        nodes = NodesDF, options = options), width = width, height = height,
        htmlwidgets::sizingPolicy(padding = 10, browser.fill = TRUE),
        package = "networkD3")
}

#' @rdname networkD3-shiny
#' @export
sankeyNetworkOutput <- function(outputId, width = "100%", height = "500px") {
    shinyWidgetOutput(outputId, "sankeyNetwork", width, height,
        package = "networkD3")
}

#' @rdname networkD3-shiny
#' @export
renderSankeyNetwork <- function(expr, env = parent.frame(), quoted = FALSE) {
    if (!quoted)
        {
            expr <- substitute(expr)
        }  # force quoted
    shinyRenderWidget(expr, sankeyNetworkOutput, env, quoted = TRUE)
}

# Need some utils functions to make this all work 
# https://github.com/christophergandrud/networkD3/blob/master/R/utils.R
check_zero <- function(Source, Target) {
    if (!is.factor(Source) && !is.factor(Target)) {
        SourceTarget <- c(Source, Target)
        if (is.numeric(SourceTarget) | is.integer(SourceTarget)) {
            if (!(0 %in% SourceTarget))
                warning(
                    'It looks like Source/Target is not zero-indexed. This is required in JavaScript and so your plot may not render.',
                    call. = FALSE)
        }
    }
}


tbl_df_strip <- function(x) {
    if('tbl_df' %in% class(x)) {
        message(paste(deparse(substitute(x)),
                      'is a tbl_df. Converting to a plain data frame.'))
        x <- base::as.data.frame(x)
    }
    return(x)
}

margin_handler <- function(margin){
  # margin can be either a single value or a list with any of
  #    top, right, bottom, left
  # if margin is a single value, then we will stick
  #    with the original behavior of networkD3 and use it for the right margin
  if(!is.null(margin) && length(margin) == 1 && is.null(names(margin))){
    margin <- list(
      top = NULL,
      right = margin,
      bottom = NULL,
      left = NULL
    )
  } else if(!is.null(margin)){
    # if margin is a named vector then convert to list
    if(!is.list(margin) && !is.null(names(margin))){
      margin <- as.list(margin)
    }
    # if we are here then margin should be a list and
    #   we will use the values supplied with NULL as default
    margin <- modifyList(
      list(top = NULL, right = NULL, bottom = NULL, left = NULL),
      margin
    )
  } else {
    # if margin is null, then make it a list of nulls for each position
    margin <- list(top = NULL, right = NULL, bottom = NULL, left = NULL)
  }
}
```

```{r}
URL <-"https://raw.githubusercontent.com/christophergandrud/networkD3/master/JSONdata/energy.json"

Energy <- jsonlite::fromJSON(URL)
 
# Now we have 2 data frames: a 'links' data frame with 3 columns (from, to, value), and a 'nodes' data frame that gives the name of each node.

# Thus we can plot it
sankeyNetwork1(Links = Energy$links, 
              Nodes = Energy$nodes, 
              Source = "source",
              Target = "target", 
              Value = "value", 
              NodeID = "name",
              units = "TWh", 
              fontSize = 12, 
              nodeWidth = 30)
```

```{r}
sankeyNetwork_dye <- function(Links, Nodes, Source, Target, Value,
    NodeID, NodeGroup = NodeID, LinkGroup = NULL, units = "",
    colourScale = JS("d3.scaleOrdinal(d3.schemeCategory20);"), fontSize = 7,
    fontFamily = NULL, nodeWidth = 15, nodePadding = 10, margin = NULL,
    height = NULL, width = NULL, iterations = 32, sinksRight = TRUE)
{
    # Check if data is zero indexed
    check_zero(Links[, Source], Links[, Target])

    # Hack for UI consistency. Think of improving.
    colourScale <- as.character(colourScale)

    # If tbl_df convert to plain data.frame
    Links <- tbl_df_strip(Links)
    Nodes <- tbl_df_strip(Nodes)

    # Subset data frames for network graph
    if (!is.data.frame(Links)) {
        stop("Links must be a data frame class object.")
    }
    if (!is.data.frame(Nodes)) {
        stop("Nodes must be a data frame class object.")
    }
    # if Source or Target are missing assume Source is the first
    # column Target is the second column
    if (missing(Source))
        Source = 1
    if (missing(Target))
        Target = 2

    if (missing(Value)) {
        LinksDF <- data.frame(Links[, Source], Links[, Target])
        names(LinksDF) <- c("source", "target")
    } else if (!missing(Value)) {
        LinksDF <- data.frame(Links[, Source], Links[, Target],
            Links[, Value])
        names(LinksDF) <- c("source", "target", "value")
    }

    # if NodeID is missing assume NodeID is the first column
    if (missing(NodeID))
        NodeID = 1
    NodesDF <- data.frame(Nodes[, NodeID])
    names(NodesDF) <- c("name")

    # add node group if specified
    if (is.character(NodeGroup)) {
        NodesDF$group <- Nodes[, NodeGroup]
    }

    if (is.character(LinkGroup)) {
        LinksDF$group <- Links[, LinkGroup]
    }

    margin <- margin_handler(margin)

    # create options
    options = list(NodeID = NodeID, NodeGroup = NodeGroup, LinkGroup = LinkGroup,
        colourScale = colourScale, fontSize = fontSize, fontFamily = fontFamily,
        nodeWidth = nodeWidth, nodePadding = nodePadding, units = units,
        margin = margin, iterations = iterations, sinksRight = sinksRight) 
    # glimpse(options)
                   #    List of 12
                   # $ NodeID     : chr "name"
                   # $ NodeGroup  : chr "name"
                   # $ LinkGroup  : NULL
                   # $ colourScale: chr "d3.scaleOrdinal(d3.schemeCategory20);"
                   # $ fontSize   : num 12
                   # $ fontFamily : NULL
                   # $ nodeWidth  : num 30
                   # $ nodePadding: num 10
                   # $ units      : chr "TWh"
                   # $ margin     :List of 4
                   #  ..$ top   : NULL
                   #  ..$ right : NULL
                   #  ..$ bottom: NULL
                   #  ..$ left  : NULL
                   # $ iterations : num 32
                   # $ sinksRight : logi TRUE
# 
#     # create widget
    htmlwidgets::createWidget(name = "sankeyNetwork", 
                              x = list(links = LinksDF,
                                        nodes = NodesDF, 
                                        options = options), 
                              width = width, 
                              height = height,
                              
        htmlwidgets::sizingPolicy(padding = 10, browser.fill = TRUE),
        package = "networkD3")
}

out <- sankeyNetwork_dye(Links = Energy$links, 
              Nodes = Energy$nodes, 
              Source = "source",
              Target = "target", 
              Value = "value", 
              NodeID = "name",
              units = "TWh", 
              fontSize = 12, 
              nodeWidth = 30)

glimpse(out)
```

```{r}
df_test <- read_csv("Exp0863--20200807_sankey_dye_data.csv")
df_test


exit_values <- read_csv("Exp0863--20200807_exit_point_values.csv")

hit_rate <- readRDS("Exp0810--20200527_dye_hit_rate_after_recall.rds")
screen_outcome <- read_csv("Exp0810--20200416_all_screen_hits_v5.csv")

df <- hit_rate %>%
  left_join(screen_outcome %>% select(dye, detect_ever, reject_buffer)) %>%
  left_join(df_test %>% select(Class, dye, start, start_join,full_lib,  exit_point, exit_reason)) %>%
  arrange(Class, dye_order_num) %>%
  mutate(Class_order = c(1:nrow(.))) %>%
  mutate_at("start", as.numeric) %>%
  mutate_at("start_join", as.numeric)

df
```
```{r}
df_long <- df %>%
  pivot_longer(cols = c(start, start_join, full_lib), names_to = "check_point", values_to = "xvalue") %>%
  group_by(xvalue, check_point)%>%
  mutate(yvalue = row_number())

df

df_long1 <- df  %>%
         pivot_longer(cols = c(start, start_join), names_to = "check_point", values_to = "xvalue") %>%
          filter(check_point == "start") %>%
          mutate(yvalue = Class_order)

df_long2 <- df  %>%
         pivot_longer(cols = c(start, start_join), names_to = "check_point", values_to = "xvalue") %>%
          filter(check_point == "start_join") %>%
          mutate(yvalue = dye_order_num)

df_long3 <- df  %>%
         pivot_longer(cols = c(start, start_join), names_to = "check_point", values_to = "xvalue") %>%
          filter(check_point == "start_join") %>%
          mutate(yvalue = dye_order_num)

full_lib
# %>%
  # pivot_longer(cols = c(Class_order, dye_order_num), names_to =  "yvar", values_to = "yvalue")

df_long <- bind_rows(df_long1, df_long2) %>%
  group_by( xvalue, check_point) %>%
  arrange(check_point, dye_order_num) %>%
  mutate(moving_y_val = row_number())

df_long %>%
  filter(xvalue == 0.9)
```


```{r}
df_long %>%
  filter(detect_ever == 1) %>%
  ggplot(aes(x = xvalue, y = yvalue,  color = dye_order_num, group = dye)) + #
  geom_line(data = df_long, aes(x = xvalue, y = yvalue, group = dye), color = "lightgrey") +
  geom_line() +
  # geom_point() +
  # geom_point(data = df %>% filter(detect_ever == 1), shape = 1, size = 2) +
  scale_color_viridis_c() 
```







Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

