---
title: "Dye Sankey Diagram"
output: html_notebook
---

This is an R notebook dedicated to figuring out the best way to make a Sankey diagram for the dye paper. 

Will explore at least: 
- networkD3 (typical sankey package)
- ggplot with geom_smooth and scale_color_viridis_d 

```{r}
library(tidyverse)

make_node <- function( df, 
                       filter_var = NULL, 
                       node_group = NULL, 
                       node_name, 
                       xvalue_in, 
                       distinct_var # variable to make distinct within the node (useful if an entity is included in multiple nodes)
                       ){
  
                    filter_call <- rlang::parse_expr(filter_var)
                    distinct_call <- rlang::parse_expr(distinct_var)
                    
                    df %>% 
                        ungroup() %>%
                        filter(!!filter_call) %>%
                        distinct(!!distinct_call, .keep_all = TRUE) %>%
                        mutate(node_group = node_group,
                               node_name = node_name,
                               xvalue = xvalue_in) 
                          }

make_node_map <- function(df, var_list) {
                                make_node( df, 
                                           filter_var = var_list$filter_var, 
                                           node_group = var_list$node_group,
                                           node_name = var_list$node_name, 
                                           xvalue_in = var_list$xvalue_in, 
                                           distinct_var = "dye")
                              }



ggsankey <- function( nodes, 
                      x_variable = xvalue_final, 
                      y_variable = yvalue_final,
                      group_variable = dye,
                      color_variable = dye_order_num,
                      color_filter = "detect_ever == 1 & reject_buffer != 1"
                      ) {

  x_var <- enquo(x_variable)
  y_var <- enquo(y_variable)
  group_var <- enquo(group_variable)
  color_var <- enquo(color_variable)
  color_filter <- rlang::parse_expr(color_filter)
  
  nodes %>%
      ungroup() %>%
     filter(!!color_filter) %>%
    
      ggplot(     ) + 
      geom_line(data = nodes,
                 aes(x = !!x_var, y = !!y_var, group = !!group_var), size = 0.2, alpha = 0.7, color = "#d9d9d9") +

    geom_line(data = nodes %>% filter(reject_buffer == 1),
                aes(x = !!x_var, y = !!y_var, group = !!group_var), size = 0.2, alpha = 0.7, color = "#525252") +
      
    geom_line( aes(x = !!x_var, y = !!y_var,  color = !!color_var , group = !!group_var), 
                 size = 0.3, alpha = 1) +
    
    scale_color_viridis_c() 
}

nudge_node <- function(df, 
                       .corr_col, 
                       .filter_col,  
                       filt_var, 
                       rep_var,
                       overwrite = FALSE ) {
  
  #  https://adv-r.hadley.nz/evaluation.html#tidy-evaluation
  #  https://community.rstudio.com/t/using-mutate-in-a-function-with-the-new-colname-as-an-argument/10399/2
  #  https://dplyr.tidyverse.org/reference/recode.html
  
  new_corr_name <- enquo(.corr_col) %>%
                    quo_name()

  df <- df %>% 
          mutate(!!new_corr_name := replace( {{.corr_col}}, {{.filter_col}} == !!filt_var, {{.corr_col}}[{{.filter_col}} == !!filt_var] + rep_var))
  
  if (overwrite == TRUE) {
    print("overwriting")
      df <- df %>% 
           mutate(!!new_corr_name := replace( {{.corr_col}}, {{.filter_col}} == !!filt_var, rep_var))
      }

  df 
}
```

Read in the raw data
```{r}
# the initial df must contain all varaibles necessary to filter to create the nodes for the Sankey
df <- readRDS("Exp0810--20200527_dye_hit_rate_after_recall.rds") %>% # dye name, tanimoto order, and and hit rate data
      left_join(read_csv("Exp0810--20200416_all_screen_hits_v5.csv") %>% select(dye, detect_ever, reject_buffer)) %>% # manually curated triage calls
      left_join(read_csv("Exp0863--20200807_sankey_dye_data.csv") %>% select(Class, dye)) %>% # just classes associated with the dye names (A, C, T ...)
      arrange(Class, dye_order_num) 

# read create nodes from the input data and a list of nodes 
nodes <- read_csv("Exp0863--20200807_nodes.csv") %>%
          pivot_longer(cols= c(xvalue_var, xvalue_var2), names_to = "node_end", values_to = "xvalue_in")%>%
          split( . , seq(nrow( . ))) %>%
          lapply(as.list) %>%
  
          lapply( . , make_node_map, df = df) %>% # make the node map for each input node info list
          bind_rows() %>% # bind together into a single tibble
  
          group_by(xvalue) %>% # these groupings create the inter-node order AND quantitative heights
          arrange(dye_order_num) %>%
          mutate(yvalue = row_number(),
                 y_correction = 0,
                 x_correction = 0) %>%
          mutate(yvalue_center = median(yvalue)) %>%
          mutate(yvalue_final = yvalue + y_correction - yvalue_center) %>%
          mutate(xvalue_final = xvalue + x_correction) %>%
         ungroup() %>%
        arrange(node_name) %>%
        mutate(node_side = rep(c("node_start", "node_end"), times = nrow(.)/2)) %>%
        unite("node_group_side", c("node_group", "node_side"), sep = "_", remove = FALSE ) %>%
        unite("node_name_side", c("node_name", "node_side"), sep = "_", remove = FALSE ) %>%
        group_by(node_name_side, xvalue) %>%
        arrange(xvalue) %>%
        nest() %>%
        ungroup() %>%
        mutate(node_number = row_number()) %>%
        unnest(data) %>%
        mutate(sankey_right_half = node_number > 27)

```

```{r}
nudged_nodes <- nodes %>% 
          nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "start_node_end", rep_var = 0.5, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "t_sens_node_end", rep_var = 0.5, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "never_hit_node_end", rep_var = 0.8, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "library_node_end", rep_var = +1, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_1_node_start", rep_var = 0, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_1_node_end", rep_var = -0.5, overwrite  = FALSE)  %>%
          
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_2_node_start", rep_var = -3, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_2_node_end", rep_var = -2.7, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_3_node_start", rep_var = -3, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_3_node_end", rep_var = -2, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_4_node_start", rep_var = -1, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_4_node_end", rep_var = -1, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_5_node_start", rep_var = -1, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_5_node_end", rep_var = -1, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_6_node_start", rep_var = -0.5, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_6_node_end", rep_var = -0.6, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_7_node_start", rep_var = -2.2, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_7_node_end", rep_var = 3, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "ever_hit_node_start", rep_var = 3, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "ever_hit_node_end", rep_var = 4, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "start", rep_var = -80, overwrite  = FALSE)  %>% 
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "start", rep_var = -4, overwrite  = FALSE)  %>% 
            
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_T", rep_var = 220, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_T", rep_var = 2, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_C", rep_var = 245, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_C", rep_var = -0.5, overwrite  = FALSE)  %>%
          
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_A", rep_var = 315, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_A", rep_var = -0.7, overwrite  = FALSE)  %>%
          
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_L", rep_var = 335, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_L", rep_var = -2.5, overwrite  = FALSE)  %>%
            
          
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_MWA", rep_var = +75, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_MWA", rep_var = -0.5, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_SYPROlog", rep_var = 5, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_SYPROlog", rep_var = 0.5, overwrite  = FALSE)  %>%
            
          nudge_node( . , .filter_col = node_name, .corr_col = yvalue_final, filt_var = "add_MWC", rep_var = -30, overwrite  = FALSE)  %>%
          nudge_node( . , .filter_col = node_name, .corr_col = xvalue_final, filt_var = "add_MWC", rep_var = -3.5, overwrite  = FALSE)  %>%
             
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "library", rep_var = 100, overwrite  = FALSE)  %>% 
            
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "ever_hit", rep_var = 100, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "ever_hit", rep_var = 0, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "ever_hit_node_end", rep_var = 0.5, overwrite  = FALSE) %>%
            
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "never_hit", rep_var = - 100, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "never_hit", rep_var = 6, overwrite  = FALSE) %>%
            
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "t_sens", rep_var = -10, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "t_sens", rep_var = 6.5, overwrite  = FALSE) %>%
            
          
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "tech1", rep_var = 330, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "tech1", rep_var = 8, overwrite  = FALSE) %>%
            
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "tech2", rep_var = 190, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "tech2", rep_var = 8, overwrite  = FALSE) %>%
            
          nudge_node( . , .filter_col = node_group, .corr_col = yvalue_final, filt_var = "tech3", rep_var = 150, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_group, .corr_col = xvalue_final, filt_var = "tech3", rep_var = 4.2, overwrite  = FALSE) %>%
            
          nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "tech3_node_end", rep_var = 1.5, overwrite  = FALSE) %>%
          #nudge_node( . , .filter_col = node_group_side, .corr_col = xvalue_final, filt_var = "tech3", rep_var = 4.5, overwrite  = FALSE) %>%
            
            
          nudge_node( . , .filter_col = sankey_right_half, .corr_col = xvalue_final, filt_var = TRUE, rep_var = 2.5, overwrite  = FALSE) %>%
          nudge_node( . , .filter_col = node_name_side, .corr_col = xvalue_final, filt_var = "library_7_node_end", rep_var = -1, overwrite  = FALSE) 

# write_rds(nudged_nodes, "Exp0863--20200808_nudged_nodes.rds")
```


Visualize and nudge nodes
```{r}
 nudged_nodes  %>% 

  ggsankey( . , xvalue_final, yvalue_final) +
  theme_void() +
  theme(legend.position = "none")-> p

ggsave("Exp0863_test_sankey2.pdf", width = 10, height = 3)
```

```{r}
ggsankey_bw <- function( nodes, 
                      x_variable = xvalue_final, 
                      y_variable = yvalue_final,
                      group_variable = dye,
                      color_variable = dye_order_num,
                      color_filter = "detect_ever == 1 & reject_buffer != 1",
                      color_plot = TRUE,
                      color_back = "#d9d9d9",
                      color_front = "black"
                      ) {

  x_var <- enquo(x_variable)
  y_var <- enquo(y_variable)
  group_var <- enquo(group_variable)
  color_var <- enquo(color_variable)
  color_filter <- rlang::parse_expr(color_filter)
  
  nodes %>%
      ungroup() %>%
     filter(!!color_filter) %>%
    
 ggplot(     ) + 
      geom_line(data = nodes,
                 aes(x = !!x_var, y = !!y_var, group = !!group_var), size = 0.2, alpha = 0.7, color = color_back) -> p
  
  if (color_plot == TRUE) {
    p +    geom_line( aes(x = !!x_var, y = !!y_var,  color = !!color_var , group = !!group_var),
                 size = 0.3, alpha = 1) +
    scale_color_viridis_c()
  } else {

    p + geom_line( aes(x = !!x_var, y = !!y_var,  color = !!color_var , group = !!group_var),
                 size = 0.3, alpha = 1, color = color_front)
  }

}

 nudged_nodes  %>% 
  ggsankey_bw( . , 
            color_filter = "Class == 'T'",
            xvalue_final, yvalue_final,
            color_plot = FALSE,
            #  color_back = "red",
            # color_front = "blue"
            ) +
            theme_void() +
            theme(legend.position = "none")
 

 nudged_nodes  %>% 
  ggsankey_bw( . , 
            color_filter = "Class == 'T'",
            xvalue_final, yvalue_final,
            color_plot = TRUE,
            #  color_back = "red",
            # color_front = "blue"
            ) +
            theme_void() +
            theme(legend.position = "none")
 
  nudged_nodes  %>% 
  ggsankey_bw( . , 
            color_filter = "dye == 'T004'",
            xvalue_final, yvalue_final,
            color_plot = FALSE,
            #  color_back = "red",
            # color_front = "blue"
            ) +
            theme_void() +
            theme(legend.position = "none")

```

```{r}
ggsankey_all_color <- function( nodes, 
                      x_variable = xvalue_final, 
                      y_variable = yvalue_final,
                      group_variable = dye,
                      color_variable = dye_order_num,
                      color_filter = "detect_ever == 1 & reject_buffer != 1"
                      ) {

  x_var <- enquo(x_variable)
  y_var <- enquo(y_variable)
  group_var <- enquo(group_variable)
  color_var <- enquo(color_variable)
  color_filter <- rlang::parse_expr(color_filter)
  
  nodes %>%
      ungroup() %>%
    # filter(!!color_filter) %>%
    
      ggplot(     ) + 
    # 
    #   geom_line(data = nodes,
    #              aes(x = !!x_var, y = !!y_var, group = !!group_var), size = 0.2, alpha = 0.7, color = "#d9d9d9") +
    # 
    # geom_line(data = nodes %>% filter(reject_buffer == 1),
    #             aes(x = !!x_var, y = !!y_var, group = !!group_var), size = 0.2, alpha = 0.7, color = "#525252") +
      
    geom_line( aes(x = !!x_var, y = !!y_var,  color = !!color_var , group = !!group_var), 
                 size = 0.3, alpha = 1) +
    
    scale_color_viridis_c() 
}

  nudged_nodes  %>% 
  ggsankey_all_color ( . , 
            color_filter = "dye == 'T004'",
            xvalue_final, yvalue_final,
            #color_plot = FALSE,
            #  color_back = "red",
            # color_front = "blue"
            ) +
            theme_void() +
            theme(legend.position = "none")
```











Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

